---
title: "Credit Card Default"
Author: Doudou Zhang
output:
  word_document:
    toc: yes
  html_notebook:
    number_sections: yes
    toc: yes
---

# Intro
As always, let's start with clearing the workspace and load required packages. 
For this exercise, we will use the dataset, *Default*, available from *ISLR* library. 
```{r}
rm(list = ls())     # clear the workspace 
library(ISLR)       # load ISLR data package
library(tidyverse)
library(ggplot2)
```


# Data Preparation
Examine the dataset. It contains four variables, `default`, `student`,`balance`, and `income`.  
```{r}
Default<-as_tibble(Default)
Default
glimpse(Default)       

head(Default) # show the first six rows
tail(Default) # show the last six rows
names(Default)      # variable names
nrow(Default)       # the number of rows
ncol(Default)       # the number of columns
summary(Default)    # basic summary statistics of the variables in Default dataset (default, student, balance, income)
```

```{r}
# frequency table 
summary(Default$default)   # summary of default variable
table(Default$default)     # contingency table: frequency of each case (yes/no) in default variable
table(Default$student)     # contingency table: frequency of each case (yes/no) in student variable
table(Default$default, Default$student)  # cross-tabulation
```

# Data Visualization
## Bar chart
```{r}
Default %>%
  ggplot(aes(x=default,fill=default)) +
  geom_bar()

Default %>%
  ggplot(aes(x=student,fill=student)) +
  geom_bar()
```

## Histograms
```{r}
Default %>%
  ggplot(aes(x=income)) +
  geom_histogram(binwidth=1000, colour="black",fill="white")

Default %>%
  ggplot(aes(x=income,fill=student)) +
  geom_histogram(binwidth=1000,alpha=.5,position="identity")

Default %>%
  ggplot(aes(x=income,fill=default)) +
  geom_histogram(binwidth=1000,alpha=.5,position="identity")
```

## Boxplots
```{r}
ggplot(Default,aes(x=default,y=balance,fill=default))+geom_boxplot()
ggplot(Default,aes(x=default,y=income,fill=default))+geom_boxplot()
```

## Scatter plots
```{r}
Default %>%
  ggplot(aes(x=balance,y=income,color=default)) +
  geom_point(shape=1)
```



## Classification Trees
We need *rpart* package for Classification Trees: http://www.statmethods.net/advstats/cart.html
`rpart.plot` is a package for visualizing classification tree models. 
```{r}
library(rpart)
library(rpart.plot)
```

Let's build a model. 
```{r}
ct_model<-rpart(default~.,           # model formula
                data=Default,                     # dataset
                method="class",                   # "class" indicates a classification tree model 
                control=rpart.control(cp=0.03,minsplit=20,maxdepth=3))   # tree control parameters. 
```

* minsplit: minimum number of data points required to attempt a split
* cp: complexity parameter
* maxdepth: depth of a classification tree 

Next, let's visualize the tree model. 
```{r}
rpart.plot(ct_model)   # tree plot
```

Here, I tried to visualize the classification tree model results on the scatterplot. It could be done, because the model used the two numeric variables. 
```{r}
Default %>%
  ggplot(aes(x=balance,y=income,color=default)) +
  geom_point(shape=1)+
  geom_vline(xintercept=1800.002,linetype="dashed")+
  geom_vline(xintercept=1971.915,linetype="dashed")+
  geom_hline(yintercept=27401.2,linetype="dashed")+
  annotate("rect",xmin=1800.002, xmax=1971.915, ymin=0, ymax=27401.2,fill="red",alpha=0.2)+
  annotate("rect",xmin=1971.915, xmax=Inf, ymin=0, ymax=Inf,fill="blue",alpha=0.2)+
  annotate("rect",xmin=0, xmax=1800.002, ymin=0, ymax=Inf,fill="red",alpha=0.2)+
  annotate("rect",xmin=1800.002, xmax=1971.915, ymin=27401.2, ymax=Inf,fill="blue",alpha=0.2)
```

```{r}
print(ct_model)        # model results 
```
```{r}
summary(ct_model)      # model result details 
```

Get the predicted value - class membership (yes or no) --> using a cut-off of 50%. 
```{r}
ct_pred_class<-predict(ct_model,type="class") # class membership (yes or no) 
head(ct_pred_class)

ct_pred<-predict(ct_model)  # get the predicted values - class probabilities (default)
head(ct_pred)
```

Let's create a new column in Default: save the predicted probability of default (yes) from the second column of dt_pred.
```{r}
Default$ct_pred_prob<-ct_pred[,2]   
```

Alternatively, you can specify a certain cut-off value to assign class membership. You can set the cut-off at 30%, 50%, 80%, or whatever you want. 
```{r}
Default$ct_pred_class<-ifelse(Default$ct_pred_prob>0.3,"Yes","No")   
```

                                                                    
```{r}
head(Default)
Default[253,]     # get the information of 253th customer 

# show the customers whose predicted probability is greater than 70%
Default%>%
  filter(ct_pred_prob>0.7) 

# sort customers by probability of default in descending order
Default%>%
  arrange(desc(ct_pred_prob))
```

## Random Forest
```{r}
#install.packages("randomForest")
library(randomForest)
rf_model<-randomForest(default~income+balance+student,              # model formula
                       data=Default)
```

```{r}
print(rf_model)
head(rf_model$votes)       # indicates the % of trees that voted for each class
head(rf_model$predicted)   # the class favored by more trees (i.e. majority vote wins) 
```

```{r}
varImpPlot(rf_model)  # importance of variables 
```
```{r}

rf_model$vote

Default$rf_vote<-predict(rf_model,Default,type="prob")[,2]
head(Default)
```

